---
title: "Data cleaning"
format: html
editor: source
---

# Setup

```{r}
#| label: load-libs
#| message: false
#| warning: false
#| output: false

library(here)
library(tidyverse)
library(janitor)
library(lubridate)
```

# Combine all yearly open pantry visits csvs into one complete file

1.  Merge the two files for 2024 into one (only once)

2.  Merge 2019 to 2025 Saturday and Friday open pantry counts (all dates)

    i)  Clean column names with `janitor::clean_names`

3.  Sort ascending by date

4.  Add whether Saturday or Friday pantry to new `weekday` column

5.  Add new `year` column for easier grouping

6.  Remove all PII

7.  Reorder columns to `visit_date`, `year`, `Weekday`, `account_number`

8.  Write the new csv file `all_openpantry_visits` to public data directory

### Merge 2024 files

2024 is split into two files because we changed pantry type from `food:open pantry` to just `open pantry` in March of 2024. The two files need to be merged and rewritten as a single csv.

Note that this only needed to be done once, it's historical data and shouldn't change.

```{r}
#| label: merge-2024
#| message: false
#| results: false

twnty24_1 <- read_csv(here("data/sensitive_data/All_Visits_by_Date-2024-1.csv"))
twnty24_2 <- read_csv(here("data/sensitive_data/All_Visits_by_Date-2024-2.csv"))

twnty24 <- bind_rows(twnty24_1, twnty24_2)

write_csv(twnty24, here("data/sensitive_data/All_Visits_by_Date-2024.csv"))
```

## Make one complete all open pantry visits file

### Load merge and sort

Load each yearly all visits file, merge them, and sort by date

```{r}
#| label: openpantry-loadmergesort
#| message: false

yrs <- c(2019, 2020, 2021, 2022, 2023, 2024, 2025)
allopenpantry_df <-
    map(yrs, function(yr) here(paste0("data/sensitive_data/All_Visits_by_Date-", yr, ".csv"))) %>% 
    map(read_csv) %>% 
    reduce(bind_rows) %>% 
    clean_names() %>% 
    arrange(visit_date)

```

### Remove PII and add weekday and year

Extract `year` and `weekday`(day of week) columns from `visit_date` using lubridate. Remove all PII and reorder columns.

```{r}
#| label: openpantry-finalize

allopenpantry_df <- allopenpantry_df %>% 
    mutate(weekday = weekdays(visit_date),
           year = year(visit_date)) %>% 
    select(c(visit_date, year, weekday, account_number))
```

### Write new file

```{r}
#| label: openpantry-writefile

allopenpantry_filepath <- here("data/all_openpantry_visits.csv")
write_csv(allopenpantry_df, allopenpantry_filepath)

```

------------------------------------------------------------------------
