---
title: "Open Pantry Timeseries"
author: "Jay Spiros"
date: "`r Sys.Date()`"
format: html
editor: source
  markdown: 
    wrap: 80
---

# Setup

```{r}
#| label: load-libs
#| message: false
#| warning: false
#| output: false

library(here)
library(tidyverse)

# variables shared by multiple notebooks, mostly filepaths
if (!exists("util_vars_loaded", mode = 'logical')) {
    source(here("src/util_vars.R"))
}
```

# Load data for the timeseries plot

```{r}
#| label: load-data
#| message: false

openpantry_df <- read_csv(here(openpantry_all_filepath))
```

The goal here is to do a linegraph of the growth over time of the Saturday and Friday pantries. It appears that we used to do the "Friday" pantry on Tuesdays before the pandemic. **NOTE**: need to ask about this. **NOTE**: Asked, we did do the "Friday" pantry on Tuesdays

```{r}
openpantry_df %>% 
    count(year, weekday, name = "visits")
#    summarize(.by = c(year, weekday), # also works, less concise
#             visits = n())

```

# Plot

For now treat Tuesday and Friday visits as "Friday" visits, and ignore visits other than Saturday and Friday visits.

The code below does the following steps:

1.  Replace "Thursday" with "Friday" in the `weekday` columns
2.  Filter out any random weekdays, keeping "Friday" and "Saturday"
3.  Add a new `visits` column, grouped by year then weekday that counts all visits for that weekday
4.  Adds a "Total" row beneath each year group with the Total visits for that year
    i)  This is done by grouping the rows per year
    ii) Using `group_modify` to bind a new row to each group with the total
    iii) Remember to ungroup afterward or you'll end up with a bunch of nested tibbles

```{r}

openpantry_visits_yearly <- 
    openpantry_df %>% 
    filter(year != year(Sys.Date())) %>% 
    mutate(weekday = replace_values(weekday, "Tuesday" ~ "Friday")) %>% 
    filter(weekday %in% c("Saturday", "Friday")) %>% 
    count(year, weekday, name = "visits") %>% 
    group_by(year) %>%
    group_modify(\(x, ...) x %>% 
        bind_rows(
            summarise(., weekday = "Total", visits = sum(visits))
        )
    ) %>% ungroup()
    
```

```{r}
head(openpantry_visits_yearly, 10)
```

```{r}
#| label: plot-year-data
#| fig-width: 12
#| fig-height: 6

ggplot(data = openpantry_visits_yearly, mapping = aes(x = factor(year), y = visits, group = weekday)) +
    geom_line(aes(color = weekday), linewidth = 1.3) +
    scale_color_manual(values = c("#FE922AFF", "#BE2102FF", "#4490FEFF" )) +
    scale_x_discrete("Year", expand = c(0,0)) + # "discrete" necessary to display every year
    theme_light() +
    labs(
        title = "Yearly Open Pantry Visits by Weekday, 2019 - 2025",
        x = "Year",
        y = "Visits"
    ) +
    theme(legend.title = element_blank())

    
```

For clarity, we'll limit the monthly data (total, not per pantry day) to a sliding 3 year window ending in the last complete month.

```{r}
start_date <- floor_date(Sys.Date(), "month") - months(1) - years(3)
end_date <- floor_date(Sys.Date(), "month") - days(1)

# floor_date below is needed to consolidate all days in a particular month into 1 date so count()
# groups the counts by month, rather than multiple days per month. 
openpantry_visits_monthly <- 
    openpantry_df %>%
    filter(between(visit_date, start_date, end_date)) %>% 
    mutate(visit_ym = floor_date(visit_date, "month"), # floor_date needed
           month_name = month(visit_date, label = TRUE)) %>% 
    count(visit_ym, month_name, name = "visits")

head(openpantry_visits_monthly)
```

```{r}
#| label: plot-month-data
#| fig-width: 10
#| fig-height: 6

# calculate start month/year string and end month/end string for the subtitle.
start_date_subtitle <- paste(month(start_date, label = TRUE), year(start_date))
end_date_subtitle <- paste(month(end_date, label = TRUE), year(end_date))

ggplot(data = openpantry_visits_monthly, mapping = aes(x = visit_ym, y = visits)) +
    geom_col(aes(fill = month_name), color = "grey40") +
    scale_x_date(date_labels = "%b-%Y", 
                 date_breaks = "1 month",
                 expand = c(0.01, 0)) + # reduces the default X axis padding
    coord_cartesian(ylim = c(250, NA)) +
    scale_fill_viridis_d(option = "turbo", direction = 1) + 
    theme_light() +
    labs(
        title = "Total Open Pantry Visits by Month",
        subtitle = str_glue("{start_date_subtitle} - {end_date_subtitle}"),
        x = "Month and Year",
        y = "Visits"
    ) +
    theme(axis.text.x = element_text(angle = 45, 
                                     size = 10, 
                                     hjust = 1.1,
                                     vjust = 1),
          legend.position = "none")
    
```

```{r}
# fully config saving plots
# ggsave(here("testplot.png"), width = 20, height = 12)
```

#### Random collection of ggplot code, saved in case it is needed later

``` R
# Code below adds points and numbers to the graph.
# disabling for now to keep the graph simple
# geom_point(aes(color = weekday), size = 3) +
# geom_text(data = openpantry_visits_yearly %>% filter(weekday == "Friday"),
#           aes(x = factor(year), y = visits, group = weekday, label = visits),
#           vjust = 1.5,
#           hjust = 0.5
#           ) +
# geom_text(data = openpantry_visits_yearly %>% filter(weekday %in% c("Saturday", "Total")),
#           aes(x = factor(year), y = visits, group = weekday, label = visits),
#           vjust = -1.2,
#           hjust = 0.5
#           ) +
# scale_color_manual(values = c("chocolate2","chartreuse4","firebrick2")) +
#scale_color_manual(values = c("#30123BFF", "#4454C4FF", "#4490FEFF", "#1FC8DEFF", "#29EFA2FF", ,"#7DFF56FF", "#C1F334FF", "#F1CA3AFF", "#FE922AFF", "#EA4F0DFF", "#BE2102FF", "#7A0403FF" ))
```
