---
title: "Open Pantry Timeseries"
author: "Jay Spiros"
date: "`r Sys.Date()`"
format: html
editor: source
---

# Setup

```{r}
#| label: load-libs
#| message: false
#| warning: false
#| output: false

library(here)
library(tidyverse)

# variables shared by multiple notebooks, mostly filepaths
if (!exists("util_vars_loaded", mode = 'logical')) {
    source(here("src/util_vars.R"))
}
```

# Load data for the timeseries plot

```{r}
#| label: load-data
#| message: false

allopenpantry_df <- read_csv(allopenpantry_filepath)
```

The goal here is to do a linegraph of the growth over time of the Saturday and Friday pantries.
It appears that we used to do the "Friday" pantry on Tuesdays before the pandemic.
**NOTE**: need to ask about this.

```{r}
allopenpantry_df %>% 
    count(year, weekday, name = "visits")
#    summarize(.by = c(year, weekday), # also works, less concise
#             visits = n())

```

# Plot

For now treat Tuesday and Friday visits as "Friday" visits, and ignore visits other than Saturday and Friday visits.

The code below does the following steps:
1. Replace "Thursday" with "Friday" in the `weekday` columns
1. Filter out any random weekdays, keeping "Friday" and "Saturday"
1. Add a new `visits` column, grouped by year then weekday that counts all visits for that weekday
1. Adds a "Total" row beneath each year group with the Total visits for that year
    i. This is done by grouping the rows per year
    i. using `group_modify` to bind a new row to each group with the total
    i. Remember to ungroup afterward or you'll end up with a bunch of nested tibbles

```{r}

openpantry_visits <- 
    allopenpantry_df %>% 
    mutate(weekday = replace_values(weekday, "Tuesday" ~ "Friday")) %>% 
    filter(weekday %in% c("Saturday", "Friday")) %>% 
    count(year, weekday, name = "visits") %>% 
    group_by(year) %>%
    group_modify(\(x, ...) x %>% 
        bind_rows(
            summarise(., weekday = "Total", visits = sum(visits))
        )
    ) %>% ungroup()
    
```
```{r}
head(openpantry_visits, 10)
```


```{r}
#| label: plot-year-data
#| fig-width: 12
#| fig-height: 6

ggplot(data = openpantry_visits, mapping = aes(x = factor(year), y = visits, group = weekday)) +
    geom_line(aes(color = weekday), linewidth = 1.3) +
    # Code below adds points and numbers to the graph.
    # disabling for now to keep the graph simple
    # geom_point(aes(color = weekday), size = 3) +
    # geom_text(data = openpantry_visits %>% filter(weekday == "Friday"),
    #           aes(x = factor(year), y = visits, group = weekday, label = visits),
    #           vjust = 1.5,
    #           hjust = 0.5
    #           ) +
    # geom_text(data = openpantry_visits %>% filter(weekday %in% c("Saturday", "Total")),
    #           aes(x = factor(year), y = visits, group = weekday, label = visits),
    #           vjust = -1.2,
    #           hjust = 0.5
    #           ) +
    scale_color_manual(values = c("chocolate2","chartreuse4","firebrick2")) +
    scale_x_discrete("Year") + # "discrete" necessary to display every year
    theme_minimal() +
    labs(
        title = "Yearly Open Pantry Visits by Weekday, 2019 - 2025",
        x = "Year",
        y = "Visits"
    ) +
    theme(legend.title = element_blank())

    
```
For clarity, we'll limit the monthly data (total, not per pantry day) to 2023 onward

```{r}
# floor_date rounds each month's "name" to the 1st of the month
# does not affect the actual count
openpantry_visits_monthly <- 
    allopenpantry_df %>%
    filter(year > 2022) %>% 
    filter(weekday %in% c("Tuesday", "Friday", "Saturday")) %>%
    mutate(visit_ym = floor_date(visit_date, "month"),
           month_name = month(visit_date, label = TRUE)) %>% 
    count(visit_ym, month_name, name = "visits")

openpantry_visits_monthly
```


```{r}
#| label: plot-month-data
#| fig-width: 10
#| fig-height: 6

library(ggsci)

ggplot(data = openpantry_visits_monthly, mapping = aes(x = visit_ym, y = visits)) +
    geom_col(aes(fill = month_name), color = "grey40") +
    scale_x_date(date_labels = "%b-%Y", 
                 date_breaks = "1 month",
                 expand = c(0.01, 0)) + # reduces the default X axis padding
    coord_cartesian(ylim = c(250, NA)) +
    scale_fill_viridis_d(option = "H", direction = 1) + 
    theme_light() +
    labs(
        title = "Total Open Pantry Visits by Month",
        subtitle = "Jan 2023 - Dec 2025",
        x = "Month and Year",
        y = "Visits"
    ) +
    theme(axis.text.x = element_text(angle = 45, 
                                     size = 10, 
                                     hjust = 1.1,
                                     vjust = 1),
          legend.position = "none")
    
```
```{r}
ggsave(here("testplot.png"), width = 20, height = 12)
```


