---
title: "Open Pantry Timeseries"
author: "Jay Spiros"
date: "`r Sys.Date()`"
format: html
editor: source
---

# Setup

```{r}
#| label: load-libs
#| message: false
#| warning: false
#| output: false

library(here)
library(tidyverse)

# variables shared by multiple notebooks, mostly filepaths
if (!exists("util_vars_loaded", mode = 'logical')) {
    source(here("src/util_vars.R"))
}
```

# Load data for the timeseries plot

```{r}
#| label: load-data
#| message: false

allopenpantry_df <- read_csv(allopenpantry_filepath)
```

The goal here is to do a linegraph of the growth over time of the Saturday and Friday pantries.
It appears that we used to do the "Friday" pantry on Tuesdays before the pandemic.
**NOTE**: need to ask about this.

```{r}
allopenpantry_df %>% 
    count(year, weekday, name = "visits")
#    summarize(.by = c(year, weekday), # also works, less concise
#             foo = n())

```

# Plot

For now treat Tuesday and Friday visits as "Friday" visits, and ignore visits other than Saturday and Friday visits.

The code below does the following steps:
1. Replace "Thursday" with "Friday" in the `weekday` columns
1. Filter out any random weekdays, keeping "Friday" and "Saturday"
1. Add a new `visits` column, grouped by year then weekday that counts all visits for that weekday
1. Adds a "Total" row beneath each year group with the Total visits for that year
    i. This is done by grouping the rows per year
    i. using `group_modify` to bind a new row to each group with the total
    i. Remember to ungroup afterward or you'll end up with a bunch of nested tibbles

```{r}

openpantry_visits <- 
    allopenpantry_df %>% 
    mutate(weekday = replace_values(weekday, "Tuesday" ~ "Friday")) %>% 
    filter(weekday %in% c("Saturday", "Friday")) %>% 
    count(year, weekday, name = "visits") %>% 
    group_by(year) %>%
    group_modify(\(x, ...) x %>% 
        bind_rows(
            summarise(., weekday = "Total", visits = sum(visits))
        )
    ) %>% ungroup()
    
```
```{r}
openpantry_visits
```


```{r}
#| label: plot-data
#| fig-width: 10
#| fig-height: 6

ggplot(data = openpantry_visits, mapping = aes(x = factor(year), y = visits, group = weekday)) +
    geom_line(aes(color = weekday), linewidth = 1.3) +
    scale_color_manual(values = c("chocolate2","chartreuse4","firebrick2")) +
    scale_x_discrete("Year") +
    scale_y_continuous("Visits") +
    theme_minimal()

    
```

